Migration Experience Trajectories: A Three Mode Principle Component Analysis

### 2022-05-27 preregistration/Preregistration3MPCA.tex
updated
feedback LFB included

\documentclass[]{article}
    \usepackage{lmodern}
    \usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
  \else % if luatex or xelatex
\ifxetex
\usepackage{mathspec}
\usepackage{xltxtra,xunicode}
\else
  \usepackage{fontspec}
\fi
\defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
\newcommand{\euro}{€}
        \fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
  \usepackage{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
  \usepackage[left=2.5in,bottom=1.25in,top=1.25in,right=1in]{geometry}
  \ifxetex
\usepackage[setpagesize=false, % page size defined by xetex
            unicode=false, % unicode breaks when used with xetex
            xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
bookmarks=true,
pdfauthor={},
pdftitle={Migration Experience Trajectories: A Three Mode Principle Component Analysis},
colorlinks=true,
citecolor=blue,
urlcolor=blue,
linkcolor=magenta,
pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

  \def\affdep{{"", "", "", "", "", ""}}%
  \def\affcity{{"", "", "", "", "", ""}}%
  \preauthor{\begin{changemargin}{-8pc}{0pc} \centering\large}
  \postauthor{\end{changemargin}}
\date{20. May 2022}
\predate{\begin{changemargin}{-8pc}{0pc} \centering\large\emph}
\postdate{\end{changemargin}}
\usepackage{fancyhdr}
\pagestyle{fancy}
\setlength{\headheight}{14.0pt}
\renewcommand{\headrulewidth}{0pt}
\lhead{}
\rhead{\large\textsc{\MakeLowercase{Preregistration: Migration
Experience Trajectories}}}

@@

In essence, the novel extensive longitudinal datasets come with new
forms of heterogeneity, where we have to consider differences between
people, over time, and across variables. Yet, past analytical advances
have almost exclusively pushed for top-down inferential modeling
procedures\footnote{For example, stationary lagged regression models
  that assume stable means and variances over time (incl., vector
  \change{autoregression}{autoregressive} models, dynamic structural
  equation models, autoregressive integrated moving average models, and
  cross-lagged panel analyses) or basic trajectory models (e.g., mixed
  effects models, spline regression models, and latent growth curve
  modeling).}. And while inferential model testing is certainly
important, we still miss discussions of methods for the more fundamental
task of describing, summarizing, and understanding the developmental
data patterns.

As an example, little data has thus far investigated the development of
migration experiences and no research has assessed the co-development of
multiple experience aspects. Yet, understanding how people differ in
their migration trajectories, can be crucial in understanding adaptive
and maladaptive patterns. \remove{We we need}\change{t}{T}o identify
which variables are most important in the adaptation of migrants
\add{over time}, we need methods to
\change{identify clusters of similar individuals and developments}{break down the data heterogeneity into its core components (in terms of important variables and developments)}
and we need \change{
to assess whether such clusters differ in}{ways to identify how these core components relate to}
key adaptation markers (including, well-being, intergroup anxiety,
outgroup trust, or societal participation). There is, thus, a clear need
to assess the utility of analysis procedures focused on the description
and understanding of complex dynamical data.

In this manuscript, we aim to assess the utility of one such promising
analysis for the use of social psychological experience sampling data.
Recent studies have laid out the potential effectiveness of dimension
reduction procedures, which can address the new forms of person-,
variable-, and time point heterogeneity jointly \citep[i.e., three-mode
principle component analyses, 3MPCA; e.g.,][]{Monden2015}.

\change{We will, particularly,}{To this aim, we will} analyse the data
from three experience sampling studies, which followed the migration
experiences of recent migrants to the Netherlands.
\add{All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation). However, given that past investigations of psychological acculturation have underexplored the crucial aspect of motivational experiences}
\citep{Kreienkamp2022d}\add{, the three studies have places a particular emphasis on the needs, goals, and motives of young migrants.}

@@

\textbf{Observational Study}. Data is collected from study subjects that
are not randomly assigned to a treatment. This includes surveys, natural
experiments, and regression discontinuity designs.
\note{Note: This is part of the OSF dropdown answer and we can't change the description.}

@@

All three
studies\note{Note: I now extended the introduction of the three studies in the end of the description section. So I hope this doesn't come out of the blue as much anymore.}
used an extensive longitudinal design. Using a daily diary format, for
at least 30 days participants received a short survey twice per day (at
around 12pm and 7pm). We additionally included a longer pre- and post
measurement survey the days before and after the extensive longitudinal
data collection.

@@

The data was collected as part of a larger collaboration on daily
intergroup relations. A sub-sample of variables has recently been
accessed by the research team for an unrelated analysis
\citep{Kreienkamp2022b}. Additionally, several of the variables were
prepared for a graphical presentation of the dataset. Thus far, none of
the proposed analyses have been conducted and none of the previous
analyses have been related to dimension reductions
\change{or longitudinal trajectories}{nor have any of the past analyses analysed the longitudinal nature of the data}.

@@

\add{We collected three experience sampling studies, following the daily migration experiences of young migrants, who had recently arrived in the Netherlands. }For
all three studies, the data was collected in a three-step procedure:

@@

For Study 1 (initial preliminary study) our target sample size is a sum
of 1,000 daily diary measurements. As we expected a completion rate of
around 80\% we aimed to recruite 20 participants (20 participants X 50
measurements). We further over-sampled slightly to compensate for
potential drop outs given the length and intensity of the study.
\add{The total collected sample size was thus 23 participants.}

We then used several key relationships (relevant to the broader research
team) to estimate sufficient power for a range of different analyses
(i.e., using power simulations). Based on these power simulations, for
Studies 2 and 3, our target sample size was a sum of 4,000 daily diary
measurements. With 100\% completion rate that would be archived with 67
participants (60 daily diary responses each). Given that we expect some
incomplete daily diary measurements, we aimed to recruit 80
participants. For \change{both studies}{Studies 2 and 3,} we again aimed
to over-sampled slightly to account for expected drop outs.
\add{We were ultimately able to recruited a total of 113 particpants for Study 2 and 71 participants for Study 3.}

@@

\textbf{Key variables}\\
\add{While} most of the variables are identical across all studies,
\remove{however} some differ slightly in content or wording. However,
for each study we aimed to collect the following types of variables.

@@

\textbf{Interpretation correlation variables}\\
To interpret the \add{person-mode} components identified by the 3MPCA,
we would like to correlate the component scores with a range of person-
or time point specific variables. While the full list of variables is
available in the accompanying data sheets (see
`VariableSelectionDaily.xlsx' and `VariableSelectionPrePost.xlsx'), we
provide a few examples of the targeted categories below.

@@

\textbf{Auxiliary variables for multiple imputation}\\
For the missing data imputation we follow the procedure reported by
\citet{Monden2015}, and include any pre-, post-, or daily dairy
variables that are significantly (\(p\) \textless{} .01) and
meaningfully (\(r\) \textgreater{} .3) correlated with either the key
variables or missingness on the key variables. For an overview of the
variables that qualify as auxiliary variables see the enclosed data
sheets (see `VariableSelectionDaily.xlsx' and
`VariableSelectionPrePost.xlsx'). Additional information about the
multiple imputation procedure is also available at
\protect\hyperlink{statistical-models}{Statistical models} 2.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Mean Allport's conditions. We create a mean-averaged index of
  Allport's conditions in response to past findings indicating that the
  conditions are best conceptualized jointly and as functioning together
  rather than as fully independent factors
  \citep[p.~766]{Pettigrew2006}. Similar to past studies we thus hope to
  build a global indicator \citep[e.g., see][]{Pettigrew2006}. As with
  other indices we will ensure that the individual items indeed relate
  to a common latent construct and are meaningfully combine\add{d} in an
  index. If this is not possible we will create sub-indices and/or
  assess the impact of the conditions separately.

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    The interaction with {[}name interaction partner{]} was on equal
    footing (same status)
  \item
    {[}name interaction partner{]} shared your goal ({[}free-text entry
    interaction key need{]})
  \item
    The interaction with {[}name interaction partner{]} was cooperative
  \item
    The interaction with {[}name interaction partner{]} was voluntary
  \end{enumerate}
\item
  Mean belongingness during intergroup contact

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    I shared information about myself.
  \item
    {[}name interaction partner{]} shared information about themselves.
  \end{enumerate}
\item
  Mean alternative interaction quality definition (``Overall, the
  interaction with {[}name interaction partner{]} was \ldots{}'')

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Unpleasant to Pleasant
  \item
    Superficial to Meaningful
  \item
    Ineffective to Effective
  \item
    Unimportant to Important
  \end{enumerate}
\item
  Mood Subscales (MDMQ)

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    alertness
  \item
    calmness
  \item
    valence
  \end{enumerate}
\item
  Mean anti-social behaviors

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Put someone down
  \item
    Show little attention in someones opinion
  \item
    Demeaning remarks
  \item
    Inappropriately addressing someone
  \item
    Ignored or excluded someone
  \item
    Doubt someones judgement
  \item
    Unwanted attempts of personal matters
  \end{enumerate}
\item
  Mean pro-social behaviors

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Listen to someones problems
  \item
    Cheer someone up
  \item
    Help someone get things done
  \item
    Help someone with responsibilities
  \end{enumerate}
\end{enumerate}

Additionally, in the pre- and post-measurement surveys are several
validated scales that we will combine into their sub-scales. Indices
will be created in line with the accompanying validation literature:

\add{Added:}

@@

\fbox{\parbox{\mylength}{
  \color{editPurple}
  \textbf{Main Data Splits:}
  \begin{enumerate}
    \item \textbf{Three studies (S1, S2, S3)}
      \begin{enumerate}
        \item Combined for most power in correlations (including, correlation with study indicator)
        \item Separate in follow up analysis to ensure that we do not have a simpson's paradox
        \item Explore additional ideosyncratic variables in separated follow up
      \end{enumerate}
    \item \textbf{Two question types (interaction, no interaction)}:
      \begin{enumerate}
        \item Combined for most power in correlations (including, correlation with interaction indicator)
        \item Separate in subgroup follow up analysis to ensure that we do not have a simpson's paradox
      \end{enumerate}
    \item \textbf{Three time scales (half-day, day, week)}
      \begin{enumerate}
        \item If all variables clearly show single most variance on the same time-scale, single analysis only
        \item If multiple time-scales are possible within single analysis, single analysis only (see question box below)
        \item If variance split across multiple time scales for substantial number of variables (> 1/4 of all variables), multiple analyses for the different time scales 
      \end{enumerate}
    \item \textbf{Multiple Imputation (20 imputations)}
      \begin{enumerate}
        \item Test stability
        \item Split-half analysis
        \item Combine after analysis
      \end{enumerate}
  \end{enumerate}
  \color{black}
}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Sample Selection.} \add{For} our sample selection
  \remove{consists of three main steps. We (a) aggregate the key variables over time to archive a reasonably interpretable number of time points and the remove a first proportion of missing data. }we
  address both the selection of time points and participants. Given that
  multiple imputation procedures even work well with large proportions
  of missing data {[}assuming missingness at random;
  \citet{Madley-Dowd2019}{]}, we decided on a general criterion of less
  than 45\% missingness to balance sample size retention and bias in the
  multiple imputation model. Thus, we then \remove{(b)} select the time
  points for which we have \textbackslash change\{responses from 55\% of
  the participants\}\{less than 45\% missingness\} and \remove{(c)}
  select participants who have \textbackslash change\{responded to at
  least 55\% of the surveys\}\{less than 45\% missingness\} across the
  selected time range.
  \add{Because the studies are comparable but not identical in their sampling strategy, question selection, and question wording, we also have to consider, whether to combine the samples. }
\item
  \textbf{Time point aggregation:}
  \add{We then aggregate the key variables over time to archive a reasonably interpretable number of time points and remove a first proportion of missing data.}
  Given that little data is available on the meaningful time scales of
  the selected psychological variables, we chose to determine the
  appropriate time scales using variance decomposition \citep[e.g.,
  see][]{Ram2014}. This is to say that we create multi-level
  unconditional means models (without predictors) that include possible
  nested time scales as levels.
  \change{We, for example, assess the explained variances for measurements nested in days, nested in weeks, nested in months.}{We chose to select time scales that align with common human cycles. We thus compare the variances of bi-daily, daily, and weekly aggregations. Additional aggregations of two weeks or the full four weeks might be possible but would most likely reduce the variance too much for any meaningul further reduction during the 3MPCA.}
  We then chose the ideal time scale for each of the key variables.
\end{enumerate}

\fbox{\parbox{\mylength}{
  \color{editPurple}
  \textbf{Note:} I talked about the individual time scales with Rei. And one issue that came up was that although the time points do not have to be equidistal between measurement ocasions, each variable needs to have data at each measured time point and the time scales should be equal across variables. This would mean that we will need to have the same time scale (e.g., daily) for all variables. Given that we have now only three time scales (i.e., bi-daily, daily, weekly) we could run the 3MPCA for all three time scales and assess differences (in a multi-verse style analysis).\\
  \textbf{Question:} I am not entirely sure I fully understand yet why it is not possible to aggregate items across different time windows as long as data is available for all time points. For example, if energy levels are kept at bi-daily measurements but for well-being we calculate a daily average, we could then replace both daily time points of well-being with the daily average. I understand that this sounds a bit silly because we give up the morning/afternoon variance but it might still be useful in terms of missingess reduction. Here my confused thoughts: Does this mean that 3MPCA cannot deal with variables that change at different speeds? Is this because the time components are estimated across all variables and might be mis-specified if we have different time scales? This seems counter my current understanding of the core array (i.e., interactions between the components) where time components could interact with variable components.\\
  Long story short: Maybe we can discuss this issue again, I feel like I might have misunderstood something somewhere :-D\\
  Sorry...
  \color{black}
}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  \textbf{Multiple Imputation.} We then create 20 imputed datasets to
  perform the analyses on. As outlined by \citet{Monden2015}, we will
  use the key variables themself as well as auxiliary variables to
  impute missing values. Auxiliary variables are any pre-, post-, or
  daily dairy variables that are significantly (\(p\) \textless{} .01)
  and meaningfully (\(r\) \textgreater{} .3) correlated with (1) the key
  variables or (2) missingness on the key variables.
\item
  \textbf{Three-way ANOVA.} We assess the percentages of explained
  variance for the person-, variable-, and time aspects --- which offers
  an indication of whether a 3MPCA is useful for the dataset. Based on
  the grand mean centered variables we conduct a fixed-effects three-way
  ANOVA of the person-, variable-, and time modes as well as their
  various interactions. We are particularly interested in the highest
  order interaction term Person * Variable * Time + error, as a large
  amount of variance in this effect would speak towards the
  interdependence of the three modes.
\item
  \textbf{Data preprocessing.} For the \change{three mode }{3M}PCA, we
  center (across participants but within time point) and normalize
  (within variable but across time points) all key variables. The
  between-person centering, ensures that all variations are around the
  mean trend (which is removed by the centering). The normalization
  within variable are important for ensuring equal variances across
  variables, which ensures equal weighting of the variables in the
  3MPCA. For methodological detail see
  \protect\hyperlink{transformations}{Transformations}.
\item
  \textbf{Selection of 3MPCA model complexity} We then run the 3MPCA. We
  use a generalized scree plot to select the appropriate number of
  components for each of the three modes (i.e., person, variable, time).
  To test the stability of the solution results are compared across the
  20 imputed datasets \add{and split-half stability is assessed}.
\item
  \textbf{3MPCA model fitting.} For the selected complexities we then
  extract simple component structures (i.e., for each mode individually)
  as well as the \change{array core}{core array} (the full 3MPCA array,
  which
  \change{holds the weights of all combinations of the three modes}{specifies all combinations of the three mode components}).
  A Joint Orthomax orthogonal rotation and standard weights
  \remove{(when necessary)} are used to extract human interpretable
  component scores. This procedure is done on all 20 imputed data sets.
\item
  \textbf{Generalized Procrustes rotation.} In order to combine the
  three individual component structure from the 20 data sets, we use a
  generalized Procrustes rotation --- calculating the average of each
  compontent and core array.
\item
  \textbf{Explained variance.} As fit indices we then calculate (1) fit
  percentage of the estimated array for each imputed data set (i.e.,
  explained variance around the removed general trend) as well as (2) an
  \emph{overall} fit percentage of the estimated array (i.e., explained
  variance including the general trend). Both these fit metrices should
  give an indication of how much of the original variance the reduced
  component structure was able to capture.
\item
  \textbf{Component interpretation} To embed and interpret the resulting
  components (especially the person components), we calculate
  correlations of the component scores with key adaptation markers
  (including, well-being, anxiety, trust, or societal inclusion). If
  components remain unclear still we can additionally perform k-means
  clustering on the lower dimensional space to identify clearer groups,
  which can be compared on the adaptation markers if necessary.
\end{enumerate}

@@

We provide the main equations of the procedure below, where individual
\(i\) (\(i = 1,\ ...,\ I\)) reported
variable\change{ $k$ ($k = 1,\ ...,\ K$)}{ $x$ and its naturalized form $z$ (where $[k = 1,\ ...,\ K]$ indexes all available variables)}
at time point \(t\) (\(t = 1,\ ...,\ T\)), so that \(I\)=number of
persons, \(K\)=number of items, \(T\)=number of time points within the
dataset. For further information as well as an example illustration see
Supplemental Material 3 in \citet{Monden2015}.

>>>>

\documentclass[]{article}
    \usepackage{lmodern}
    \usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
  \else % if luatex or xelatex
\ifxetex
\usepackage{mathspec}
\usepackage{xltxtra,xunicode}
\else
  \usepackage{fontspec}
\fi
\defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
\newcommand{\euro}{€}
        \fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
  \usepackage{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
  \usepackage[left=2.5in,bottom=1.25in,top=1.25in,right=1in]{geometry}
  \ifxetex
\usepackage[setpagesize=false, % page size defined by xetex
            unicode=false, % unicode breaks when used with xetex
            xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
bookmarks=true,
pdfauthor={},
pdftitle={Migration Experience Trajectories: A Three Mode Principle Component Analysis},
colorlinks=true,
citecolor=BlueViolet,
urlcolor=BlueViolet,
linkcolor=BlueViolet,
pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

  \def\affdep{{"", "", "", "", "", ""}}%
  \def\affcity{{"", "", "", "", "", ""}}%
  \preauthor{\begin{changemargin}{-8pc}{0pc} \centering\large}
  \postauthor{\end{changemargin}}
\date{27. May 2022}
\predate{\begin{changemargin}{-8pc}{0pc} \centering\large\emph}
\postdate{\end{changemargin}}
\usepackage{fancyhdr}
\pagestyle{fancy}
\setlength{\headheight}{14.0pt}
\renewcommand{\headrulewidth}{0pt}
\lhead{}
\rhead{\large\textsc{\MakeLowercase{Preregistration: Migration
Experience Trajectories}}}

@@

In essence, the novel extensive longitudinal datasets come with new
forms of heterogeneity, where we have to consider differences between
people, over time, and across variables. Yet, past analytical advances
have almost exclusively pushed for inferential modeling
procedures\footnote{For example, stationary lagged regression models
  that assume stable means and variances over time (incl., vector
  autoregressive models, dynamic structural equation models,
  autoregressive integrated moving average models, and cross-lagged
  panel analyses) or basic trajectory models (e.g., mixed effects
  models, spline regression models, and latent growth curve modeling).}.
And while inferential model testing is certainly important, we still
miss discussions of methods for the more fundamental task of describing,
summarizing, and understanding the initial developmental data patterns.

As an example, a recent review of migration experiences has pointed out
that despite many complex and dynamic theories, investigations of
migrant adaptation have undervalued developmental data --- especially
when it comes to the more internal experiences of motivations and
emotions \citep{Kreienkamp2022d}. Yet, understanding how people differ
in their migration trajectories, can be crucial in understanding
adaptive and maladaptive patterns. To identify which variables are most
important in the adaptation of migrants over time, we need methods to
break down the data heterogeneity into its core components (in terms of
important variables and developments) and we need ways to identify how
these core components relate to key adaptation markers (including,
well-being, intergroup anxiety, outgroup trust, or societal
participation). There is, thus, a clear need to assess the utility of
analysis procedures focused on the description and understanding of
complex dynamical data.

In this manuscript, we aim to assess the utility of one such promising
analysis technique --- \textit{three-mode principle component analysis}
(3MPCA). We use 3MPCA specifically because recent studies have laid out
the potential effectiveness of dimension reduction procedures, which can
address the new forms of person-, variable-, and time point
heterogeneity jointly \citep[e.g.,][]{Monden2015}. We are among the
first to apply this analysis to experience sampling data and, to the
best of our knowledge, we are the first to decompose social
psychological experiences. This stands in stark contrast, to a renewed
recognition that social psychological phenomena unfold over time, a
rapid increase of extensive longitudinal data collections, and a growing
interest in understanding the co-development of multiple experience
aspects
\citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.

To untangle the heterogeneity in real-life migrant adaptations, we will
analyse data from three experience sampling studies, which followed the
migration experiences of recent migrants to the Netherlands. All three
studies focus on the psychological adaptation of migrants (i.e.,
psychological acculturation). However, given that past investigations of
psychological acculturation have underexplored the crucial aspect of
motivational experiences \citep{Kreienkamp2022d}, the three studies have
placed a particular emphasis on the needs, goals, and motives of young
migrants. To make full use of all three data sets and to guide future
use of dimension reduction procedures with extensive psychological data,
we will analyze the data in a descriptive and step wise manner, where we
first consider the data jointly and will then assess potential
differences between datasets, variables, and time scales.

@@

\textbf{Observational Study}. Data is collected from study subjects that
are not randomly assigned to a treatment. This includes surveys, natural
experiments, and regression discontinuity designs.

@@

All three studies used an extensive longitudinal design. Using a daily
diary format, for at least 30 days participants received a short survey
twice per day (at around 12pm and 7pm). We additionally included a
longer pre- and post measurement survey the days before and after the
extensive longitudinal data collection.

@@

The data was collected as part of a larger collaboration on daily
intergroup relations. A sub-sample of variables has recently been
accessed by the research team for an unrelated analysis
\citep{Kreienkamp2022b}. Additionally, several of the variables were
prepared for a graphical presentation of the dataset. Thus far, none of
the proposed analyses have been conducted and none of the previous
analyses have been related to dimension reductions or time-series
modeling.

@@

We collected three experience sampling studies, following the daily
migration experiences of young migrants, who had recently arrived in the
Netherlands. For all three studies, the data was collected in a
three-step procedure:

@@

For Study 1 (initial preliminary study) our target sample size is a sum
of 1,000 daily diary measurements. As we expected a completion rate of
around 80\% we aimed to recruite 20 participants (20 participants X 50
measurements). We further over-sampled slightly to compensate for
potential drop outs given the length and intensity of the study. The
total collected sample size of Study 1 was thus 1,225 survey responses
from 23 participants.

For Studies 2 and 3, our target sample size was a sum of 4,000 daily
diary measurements each. With 100\% completion rate that would be
archived with 67 participants (60 daily diary responses each). Given
that we expect some incomplete daily diary measurements, we again
over-sampled in both studies. In Study 2, we were ultimately able to
collect 4,965 survey responses from 113 participants. In Study 3 we
collected 4,107 survey responses from 71 participants.

@@

\textbf{Key variables}\\
While most of the variables are identical across all studies, some
differ slightly in content or wording. However, for each study we aimed
to collect the following types of variables.

@@

For the detailed variable selection of each analysis see
`VariableSelectionAnalyses.xlsx'.

\textbf{Interpretation correlation variables}\\
To interpret the person-mode components identified by the 3MPCA, we
would like to correlate the component scores with a range of person- or
time point specific variables. While the full list of variables is
available in the accompanying data sheets (see
`VariableSelectionAnalyses.xlsx'), we provide a few examples of the
targeted categories below.

@@

\textbf{Auxiliary variables for multiple imputation}\\
For the missing data imputation we follow the procedure reported by
\citet{Monden2015}, and include any pre-, post-, or daily dairy
variables that are significantly (\(p\) \textless{} .01) and
meaningfully correlated (\(r\) \textgreater{} .3) with either the key
variables or missingness on the key variables. For an overview of the
variables that qualify as auxiliary variables see the enclosed data
sheets (see `VariableSelectionAnalyses.xlsx'). Additional information
about the multiple imputation procedure is also available at
\protect\hyperlink{statistical-models}{Statistical models} 2.

@@

˜ 1. Mean Allport's conditions. We create a mean-averaged index of
Allport's conditions in response to past findings indicating that the
conditions are best conceptualized jointly and as functioning together
rather than as fully independent factors \citep[p.~766]{Pettigrew2006}.
Similar to past studies we thus hope to build a global indicator
\citep[e.g., see][]{Pettigrew2006}. As with other indices we will ensure
that the individual items indeed relate to a common latent construct and
are meaningfully combined in an index. If this is not possible we will
create sub-indices and/or assess the impact of the conditions
separately. a. The interaction with {[}name interaction partner{]} was
on equal footing (same status) a. {[}name interaction partner{]} shared
your goal ({[}free-text entry interaction key need{]}) a. The
interaction with {[}name interaction partner{]} was cooperative a. The
interaction with {[}name interaction partner{]} was voluntary 1. Mean
belongingness during intergroup contact a. I shared information about
myself. a. {[}name interaction partner{]} shared information about
themselves. 1. Mean alternative interaction quality definition
(``Overall, the interaction with {[}name interaction partner{]} was
\ldots{}'') a. Unpleasant to Pleasant a. Superficial to Meaningful a.
Ineffective to Effective a. Unimportant to Important 1. Mood Subscales
(MDMQ) a. alertness a. calmness a. valence 1. Mean anti-social behaviors
a. Put someone down a. Show little attention in someones opinion a.
Demeaning remarks a. Inappropriately addressing someone a. Ignored or
excluded someone a. Doubt someones judgement a. Unwanted attempts of
personal matters 1. Mean pro-social behaviors a. Listen to someones
problems a. Cheer someone up a. Help someone get things done a. Help
someone with responsibilities

Additionally, in the pre- and post-measurement surveys are several
validated scales that we will combine into their sub-scales. Indices
will be created in line with their respective validation literature. For
a full list of all variables and indices see
`VariableSelectionAnalyses.xlsx'.

\color{red}

\textbf{TO BE REMOVED ONCE 'VariableSelectionAnalyses.xlsx' IS DONE!}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Sample Selection.} For our sample selection we address both
  the selection of time points and participants. Given that multiple
  imputation procedures even work well with large proportions of missing
  data {[}assuming missingness at random; \citet{Madley-Dowd2019}{]}, we
  decided on a general criterion of less than 45\% missingness to
  balance sample size retention and bias in the multiple imputation
  model. Thus, we then select the time points for which we have less
  than 45\% missingness and select participants who have less than 45\%
  missingness across the selected time range. Because the studies are
  comparable but not identical in their sampling strategy, question
  selection, and question wording, we also have to consider, whether to
  combine the samples.
\item
  \textbf{Time point aggregation.} We then aggregate the key variables
  over time to archive a reasonably interpretable number of time points
  and remove a first proportion of missing data. Given that little data
  is available on the meaningful time scales of the selected
  psychological variables, we chose to determine the appropriate time
  scales using variance decomposition \citep[e.g., see][]{Ram2014}. This
  is to say that we create multi-level unconditional means models
  (without predictors) that include possible nested time scales as
  levels. We chose to select time scales that align with common human
  cycles. We thus compare the variances of bi-daily, daily, and weekly
  aggregations. Additional aggregations of two weeks or the full four
  weeks might be possible but would most likely reduce the variance too
  much for any meaningul further reduction during the 3MPCA. We then
  chose the ideal time scale for each of the key variables.
\item
  \textbf{Multiple Imputation.} We then create 20 imputed datasets to
  perform the analyses on. As outlined by \citet{Monden2015}, we will
  use the key variables themself as well as auxiliary variables to
  impute missing values. Auxiliary variables are any pre-, post-, or
  daily dairy variables that are significantly (\(p\) \textless{} .01)
  and meaningfully (\(r\) \textgreater{} .3) correlated with (1) the key
  variables or (2) missingness on the key variables.
\item
  \textbf{Three-way ANOVA*.} We assess the percentages of explained
  variance for the person-, variable-, and time aspects --- which offers
  an indication of whether a 3MPCA is useful for the dataset. Based on
  the grand mean centered variables we conduct a fixed-effects three-way
  ANOVA of the person-, variable-, and time modes as well as their
  various interactions. We are particularly interested in the highest
  order interaction term Person * Variable * Time + error, as a large
  amount of variance in this effect would speak towards the
  interdependence of the three modes.
\item
  \textbf{Data preprocessing*.} For the 3MPCA, we center (across
  participants but within time point) and normalize (within variable but
  across time points) all key variables. The between-person centering,
  ensures that all variations are around the mean trend (which is
  removed by the centering). The normalization within variable are
  important for ensuring equal variances across variables, which ensures
  equal weighting of the variables in the 3MPCA. For methodological
  detail see \protect\hyperlink{transformations}{Transformations}.
\item
  \textbf{Selection of 3MPCA model complexity*.} We then run the 3MPCA.
  We use a generalized scree plot to select the appropriate number of
  components for each of the three modes (i.e., person, variable, time).
  To test the stability of the solution results are compared across the
  20 imputed datasets and split-half stability is assessed.
\item
  \textbf{3MPCA model fitting*.} For the selected complexities we then
  extract simple component structures (i.e., for each mode individually)
  as well as the core array (the full 3MPCA array, which specifies all
  combinations of the three mode components). A Joint Orthomax
  orthogonal rotation and standard weights are used to extract human
  interpretable component scores. This procedure is done on all 20
  imputed data sets.
\item
  \textbf{Generalized Procrustes rotation*.} In order to combine the
  three individual component structure from the 20 data sets, we use a
  generalized Procrustes rotation --- calculating the average of each
  compontent and core array.
\item
  \textbf{Explained variance*.} As fit indices we then calculate (1) fit
  percentage of the estimated array for each imputed data set (i.e.,
  explained variance around the removed general trend) as well as (2) an
  \emph{overall} fit percentage of the estimated array (i.e., explained
  variance including the general trend). Both these fit metrices should
  give an indication of how much of the original variance the reduced
  component structure was able to capture.
\item
  \textbf{Component interpretation*.} To embed and interpret the
  resulting components (especially the person components), we calculate
  correlations of the component scores with key adaptation markers
  (including, well-being, anxiety, trust, or societal inclusion). If
  components remain unclear still we can additionally perform k-means
  clustering on the lower dimensional space to identify clearer groups,
  which can be compared on the adaptation markers if necessary.
\end{enumerate}

\textbf{*Step-wise Follow-up Analyses:}

\begin{enumerate}
  \item \textbf{Three studies (S1, S2, S3)}
    \begin{enumerate}
      \item Main Analysis: Combined for most power in correlations (including, correlation with study indicator)
      \item Separate in follow up analysis to ensure that we do not have a Simpson's paradox
      \item Explore additional idiosyncratic variables in separated follow up
    \end{enumerate}
  \item \textbf{Two question types (interaction, no interaction)}:
    \begin{enumerate}
      \item Main Analysis: Combined for most power in correlations (including, correlation with interaction indicator)
      \item Separate in subgroup follow up analysis to ensure that we do not have a Simpson's paradox
    \end{enumerate}
  \item \textbf{Three time scales (half-day, day, week)}
    \begin{enumerate}
      \item Main Analysis: If all variables clearly show single most variance on the same time-scale, single analysis only
      \item If multiple time-scales are possible within single analysis, single analysis only (see question box below)
      \item If variance split across multiple time scales for substantial number of variables (> 1/4 of all variables), multiple analyses for the different time scales 
    \end{enumerate}
\end{enumerate}

@@

We provide the main equations of the procedure below, where individual
\(i\) (\(i = 1,\ ...,\ I\)) reported variable \(x\) and its naturalized
form \(z\) (where \([k = 1,\ ...,\ K]\) indexes all available variables)
at time point \(t\) (\(t = 1,\ ...,\ T\)), so that \(I\)=number of
persons, \(K\)=number of items, \(T\)=number of time points within the
dataset. For further information as well as an example illustration see
Supplemental Material 3 in \citet{Monden2015}.
